The Talking Heads Experiment

### 2015-04-21 chap10/chap10.tex
fixed some more widows/orphans

\begin{figure}[p]
  \centerline{\includegraphics[width=.60\textwidth]{chap10/figs/PERACT-setup.pdf}}
\caption{\label{fig:PERACT-setup}The installation for the case grammar experiments used a more sophisticated vision system and objects that 
could be moved around. We see the two Talking Heads cameras in 
the background. They are oriented towards objects on the table or situations in the 
environment. The monitors show the camera images and the states of image processing, just as for the original 
Talking Heads experiment. 
}
\end{figure}

\begin{figure}[p]
  \centerline{\includegraphics[width=.80\textwidth]{chap10/figs/PERACT-images.pdf}}
\caption{\label{fig:PERACT-images}Top: Some of the images recorded during the action of picking up a cylinder. 
Middle: Results of segmentation algorithms. 
Bottom: The spatio-temporal continuity of objects is tracked. 
}
\end{figure}

>>>>

\begin{figure}[p]
  \centerline{\includegraphics[width=.60\textwidth]{chap10/figs/peract-setup.pdf}}
\caption{\label{fig:PERACT-setup}The installation for the case grammar experiments used a more sophisticated vision system and objects that 
could be moved around. We see the two Talking Heads cameras in 
the background. They are oriented towards objects on the table or situations in the 
environment. The monitors show the camera images and the states of image processing, just as for the original 
Talking Heads experiment. 
}
\end{figure}

\begin{figure}[p]
  \centerline{\includegraphics[width=.80\textwidth]{chap10/figs/peract-images.pdf}}
\caption{\label{fig:PERACT-images}Top: Some of the images recorded during the action of picking up a cylinder. 
Middle: Results of segmentation algorithms. 
Bottom: The spatio-temporal continuity of objects is tracked. 
}
\end{figure}

### 2015-04-21 chap12/chap12.tex
fixed some more widows/orphans

\begin{figure}[p]
\centerline{
\includegraphics[width=\textwidth]{chap12/figs/full-cs-graph.pdf}
}\\
\centerline{
\includegraphics[height=0.5\textwidth]{chap12/figs/full-1000.pdf}$~~~$
\includegraphics[height=0.5\textwidth]{chap12/figs/full-3000.pdf}
}
\caption{{The {\itshape Basic Colour Strategy} allows a population of agents (in this case 10) to self-organise a colour lexicon from scratch. The graph shows that high communicative success is reached with a lexicon of about 15 colour words (top). The evolution of a typical lexicon in a smaller population (5 agents), is shown after 400 (bottom left) and 1200 (bottom right) games per agent. Each row represents the lexicon of one agent.}
\label{fig:full-cs-dynamics}
} 
\end{figure}

@@

\begin{figure}[p]
\centerline{
\includegraphics[width=\textwidth]{chap12/figs/brightness-graph.pdf}
}\\
\centerline{
\includegraphics[height=0.5\textwidth]{chap12/figs/bw-1000.pdf} $~~~$
\includegraphics[height=0.5\textwidth]{chap12/figs/bw-4000.pdf}}
\label{fig:brightness-dynamics}
\caption{{The {\itshape Brightness Colour Strategy} also allows a population of agents (in this case 10)
to evolve an adequate colour lexicon (top). The evolution of a typical lexicon in a smaller population (5 agents), is shown 
after 400 (bottom left) and 1600 (bottom right) games per agent. Each row represents
the lexicon of one agent.}}
\end{figure}

@@

A {\itshape conceptualisation strategy} is an abstraction of such a concrete conceptualisation network. 
\is{conceptualisation strategy}
It consists of a network that acts as a useful chunk and is therefore a modular unit that can be linked into a larger network. 
The chunk has a set of open variables with which the network can be linked to other networks. 
For example, the network shown in \figref{fig:spatial-strategies} can be made more abstract to handle many instances 
of proximal relations, not only `near the box', but also `far from the box', `close to the box', `near the boxes', 
`near the table', `very far from the wall', etc. 
The open variables are ?cat, ?selector-2, and ?landmark-class, which have to be supplied by lexical items or 
phrases. The network is then as follows, with operations at 3, 6 and 8 removed: 
\begin{small}
\begin{verbatim}
1. (get-context ?ctx-1) 
2. (construct-region-proximal ?region ?ctx-1 ?landmark ?cat) 
4. (apply-selection ?landmark ?landmarks ?selector-2)
5. (apply-class ?landmarks ?ctx ?landmark-class) 
7. (get-context ?ctx) 
\end{verbatim}
\end{small}
The network has to be completed with bindings for the open variables, but all the other variables can be 
computed through the network itself. 
Network-chunks are expressed through a grammatical construction, which expresses which chunk is involved 
and how the initial values of the slots are to be expressed. In this 
case it is a noun-phrase construction, syntactically of the form [noun preposition article noun] and 
semantically constrained as [proximal-category selector object-class].

@@

There are also strategies that focus 
on minimising the cognitive effort of the listener by signalling how words in the sentence need to be semantically grouped together. 
This happens generally in two ways: 
\enlargethispage{1\baselineskip}\begin{itemize}
\item Language-dependent markers for various syntactico-semantic features, such as number, gender, case, are introduced 
and the different words which need to be grouped are all marked for the same features. An example of this kind of 
grammatical agreement is `la belle fille' (French: the beautiful girl) where the article `la' and 
the adjective `belle' are both marked as singular feminine. 
\item The second strategy is phrase structure. Words are grouped together that are about the same object, 
a sequential order is imposed on these words, and patterns of orderings based on syntactic and semantic categories are
introduced. 
\end{itemize}
Extensive empirical work on grammaticalisation \is{grammaticalisation}
has given us a wealth of data how these strategies play out in the history 
of human languages.\footnote{
See for example: \cite{Heine:2007}, \cite{Hopper:1993}, \cite{Lehmann:1982}, \cite{Traugott:2013}.}
And the key challenge for evolutionary linguistics is to model 
these processes, the same way evolutionary biologists model speciation or species evolution.

>>>>

\begin{figure}[p]
\centerline{
\includegraphics[width=\textwidth]{chap12/figs/full-cs-graph.pdf}
}
%\\

\centerline{
\includegraphics[height=0.5\textwidth]{chap12/figs/full-1000.pdf}$~~~$
\includegraphics[height=0.5\textwidth]{chap12/figs/full-3000.pdf}
}
\caption{{The {\itshape Basic Colour Strategy} allows a population of agents (in this case 10) to self-organise a colour lexicon from scratch. The graph shows that high communicative success is reached with a lexicon of about 15 colour words (top). The evolution of a typical lexicon in a smaller population (5 agents), is shown after 400 (bottom left) and 1200 (bottom right) games per agent. Each row represents the lexicon of one agent.}
\label{fig:full-cs-dynamics}
} 
\end{figure}

@@

\begin{figure}[p]
\centerline{
\includegraphics[width=\textwidth]{chap12/figs/brightness-graph.pdf}
}
% \\

\centerline{
\includegraphics[height=0.5\textwidth]{chap12/figs/bw-1000.pdf} $~~~$
\includegraphics[height=0.5\textwidth]{chap12/figs/bw-4000.pdf}}
\label{fig:brightness-dynamics}
\caption{{The {\itshape Brightness Colour Strategy} also allows a population of agents (in this case 10)
to evolve an adequate colour lexicon (top). The evolution of a typical lexicon in a smaller population (5 agents), is shown 
after 400 (bottom left) and 1600 (bottom right) games per agent. Each row represents
the lexicon of one agent.}}
\end{figure}

@@

A {\itshape conceptualisation strategy} is an abstraction of such a concrete conceptualisation network. 
\is{conceptualisation strategy}
It consists of a network that acts as a useful chunk and is therefore a modular unit that can be linked into a larger network. 
The chunk has a set of open variables with which the network can be linked to other networks. 
For example, the network shown in \figref{fig:spatial-strategies} can be made more abstract to handle many instances 
of proximal relations, not only `near the box', but also `far from the box', `close to the box', `near the boxes', 
`near the table', `very far from the wall', etc. 
\enlargethispage{1em}
The open variables are ?cat, ?selector-2, and ?landmark-class, which have to be supplied by lexical items or 
phrases. The network is then as follows, with operations at 3, 6 and 8 removed: 
\begin{small}
\begin{verbatim}
1. (get-context ?ctx-1) 
2. (construct-region-proximal ?region ?ctx-1 ?landmark ?cat) 
4. (apply-selection ?landmark ?landmarks ?selector-2)
5. (apply-class ?landmarks ?ctx ?landmark-class) 
7. (get-context ?ctx) 
\end{verbatim}
\end{small}
The network has to be completed with bindings for the open variables, but all the other variables can be 
computed through the network itself. 
Network-chunks are expressed through a grammatical construction, which expresses which chunk is involved 
and how the initial values of the slots are to be expressed. In this 
case it is a noun-phrase construction, syntactically of the form [noun preposition article noun] and 
semantically constrained as [proximal-category selector object-class].

@@

There are also strategies that focus 
on minimising the cognitive effort of the listener by signalling how words in the sentence need to be semantically grouped together. 
This happens generally in two ways:

\begin{itemize}
\item Language-dependent markers for various syntactico-semantic features, such as number, gender, case, are introduced 
and the different words which need to be grouped are all marked for the same features. An example of this kind of 
grammatical agreement is `la belle fille' (French: the beautiful girl) where the article `la' and 
the adjective `belle' are both marked as singular feminine. 
\item The second strategy is phrase structure. Words are grouped together that are about the same object, 
a sequential order is imposed on these words, and patterns of orderings based on syntactic and semantic categories are
introduced. 
\end{itemize}
Extensive empirical work on grammaticalisation \is{grammaticalisation}
has given us a wealth of data how these strategies play out in the history 
of human languages.\footnote{
See for example: \cite{Heine:2007}, \cite{Hopper:1993}, \cite{Lehmann:1982}, \cite{Traugott:2013}.}
And the key challenge for evolutionary linguistics is to model 
these processes, the same way evolutionary biologists model speciation or species evolution.

### 2015-04-21 chap9/chap9.tex
fixed some more widows/orphans

@@

The creation of a shared language by a group of autonomous distributed agents is extraordinarily difficult because there are many sources of noise, in the form of disturbances that cause incoherence between the agents:

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.55\textwidth]{chap9/figs/london-heads.pdf} 
\includegraphics[width=.44\textwidth]{chap9/figs/screen.pdf}}
\caption{\label{fig:london-heads}Installation at the Wellcome Gallery in London. Left: Talking Heads cameras oriented towards the wall on which 
geometric figures were pasted. Right: Projection of interaction during ongoing experiment. A game just failed and 
the speaker says ``no''.} 
\end{figure}

@@

{\bfshape 2000-02-28 17:04:23 Yeah8a8y}\\
{\itshape RE: Collusion  Tribes}\\
I'm posed with another problem... As you know Brussels and Paris have been inactive for a bit... do I keep with the project and just send the bots out to Paris... or use this as a change point and educate them in Brussel's...\\
I'm going to stay with Paris... Teach other people the use of the words carefully set out by me and Trash... and generally gather more success!\\
I now have 18 bots by the way most are ranked within the top 100.. 1 or 2 exeptions... anyway I have been successful by sending the all to Paris... for top gamage action \\
It usaully takes all night but there you go... even the Earth took an entire week to sculpture.

% \rule{0.8\textwidth}{.4pt}
\end{mail}
\todo{removed one \textbackslash rule from the last entry. Also, why is \textbackslash rule set to .8 of textwidth?!}
\enlargethispage{1\baselineskip}
To deal with the problem of agent congestion, software issues, and camera misalignment at the London 
and Cambridge sites, maintenance was carried out which lead to a temporary unavailability of the experiment. 
Surprisingly, the reaction of the Hull group was entirely negative as they saw it as a 
way to prevent them from gaining or keeping control of the language used by the agents, even though 
Angus McIntyre clearly and patiently explained why maintenance was necessary. Also, Denna Jones ({\it Phlox}) took 
it personally as she did not realise that the calibration errors were due to vibrations caused by
traffic of the London Underground near and under the Wellcome Gallery.

@@

@@

{\bfshape 2000-03-07 22:10:09 Yeah8a8y}\\
{\itshape RE: Nowhere to launch!!!}\\
All hail ``cheesyslurpscum''!!!  Very good aimed at the cheesymeister himself... ``hullsuckx'' indeed!!!\\
Chris\\
Resident Hull Uni director of Poonani

\end{mail}

@@

{\bfshape 2000-03-08 16:01:19 Trash}\\
{\itshape Ahhh thats why!!}\\
Oh nice to see that ...  I thought that you just hated asomeone caled adam wasn't aware that adam IS A CUNT... 
may have to reopinionate myself with you ...  nice on efella you deserve respect \\
Hull Uni Coordionator of Total System Breakdown

\end{mail}

@@

{\bfshape 2000-03-09 16:13:23 phlox}\\
{\itshape we're still waiting. . .}\\
come on sony own up. if youre gonna exercise stalinistic control at least do it up front; not under the guise of testing software;. the only agents on site at the mo are owned by luc, angus, adam and joris. if it's in the public domain - and it is -  then let the public play!

\end{mail}
\clearpage
The postings on the Forum keep going in crescendo for a while until the exhibition ends March 2000.

>>>>

\newpage

@@

\enlargethispage{1em}
The creation of a shared language by a group of autonomous distributed agents is extraordinarily difficult because there are many sources of noise, in the form of disturbances that cause incoherence between the agents:

@@

\begin{figure}[htbp]
 \centerline{
  \includegraphics[width=.55\textwidth]{chap9/figs/london-heads.pdf} 
\includegraphics[width=.44\textwidth]{chap9/figs/screen.pdf}
}
\caption{Installation at the Wellcome Gallery in London. Left: Talking Heads cameras oriented towards the wall on which 
geometric figures were pasted. Right: Projection of interaction during ongoing experiment. A game just failed and 
the speaker says ``no''.} 
\label{fig:london-heads}
\end{figure}

@@

{\bfshape 2000-02-28 17:04:23 Yeah8a8y}\\
{\itshape RE: Collusion  Tribes}\\
I'm posed with another problem... As you know Brussels and Paris have been inactive for a bit... do I keep with the project and just send the bots out to Paris... or use this as a change point and educate them in Brussel's...\\
I'm going to stay with Paris... Teach other people the use of the words carefully set out by me and Trash... and generally gather more success!\\
I now have 18 bots by the way most are ranked within the top 100.. 1 or 2 exeptions... anyway I have been successful by sending the all to Paris... for top gamage action \\
It usaully takes all night but there you go... even the Earth took an entire week to sculpture.
%
% \rule{0.8\textwidth}{.4pt}
\end{mail}

\todo{removed one \textbackslash rule from the last entry. Also, why is \textbackslash rule set to .8 of textwidth?!}
\enlargethispage{1\baselineskip}

To deal with the problem of agent congestion, software issues, and camera misalignment at the London 
and Cambridge sites, maintenance was carried out which lead to a temporary unavailability of the experiment. 
Surprisingly, the reaction of the Hull group was entirely negative as they saw it as a 
way to prevent them from gaining or keeping control of the language used by the agents, even though 
Angus McIntyre clearly and patiently explained why maintenance was necessary. Also, Denna Jones ({\it Phlox}) took 
\enlargethispage{1em}
it personally as she did not realise that the calibration errors were due to vibrations caused by
traffic of the London Underground near and under the Wellcome Gallery.

@@

\newpage
\begin{mail}

@@

{\bfshape 2000-03-07 22:10:09 Yeah8a8y}\\
{\itshape RE: Nowhere to launch!!!}\\
All hail ``cheesyslurpscum''!!!  Very good aimed at the cheesymeister himself... ``hullsuckx'' indeed!!!\\
Chris\\
Resident Hull Uni director of Poonani
%
\end{mail}

@@

{\bfshape 2000-03-08 16:01:19 Trash}\\
{\itshape Ahhh thats why!!}\\
Oh nice to see that ...  I thought that you just hated asomeone caled adam wasn't aware that adam IS A CUNT... 
may have to reopinionate myself with you ...  nice on efella you deserve respect \\
Hull Uni Coordionator of Total System Breakdown
%
\end{mail}

@@

{\bfshape 2000-03-09 16:13:23 phlox}\\
{\itshape we're still waiting. . .}\\
come on sony own up. if youre gonna exercise stalinistic control at least do it up front; not under the guise of testing software;. the only agents on site at the mo are owned by luc, angus, adam and joris. if it's in the public domain - and it is -  then let the public play!
%
\end{mail}
\clearpage

The postings on the Forum keep going in crescendo for a while until the exhibition ends March 2000.
